Active Learning from Blackbox to Timed Connectors

### 2016-01-15 paper/paper.tex
.

% FIXME finally no todos should exist in this draft
\usepackage[textsize=tiny]{todonotes}

@@

Distributed real-time embedded system (DRES) is reforming our lives with the name \emph{IoT}, the
internet of things, wherein systems are usually composed of individual components and a middleware
serving as a \emph{connector}. Such systems could be distributed logically or physically, which
makes the coordination even more complicated. In this case, we need to specify these coordination
processes with so-called \emph{coordination languages}, so that formal techniques can be applied to
guarantee their reliability.

\emph{Timed Reo} is a real-time extension of the coordination language \emph{Reo}. With timeline
involved, coordination process in DRES can be depicted clearly and intuitively. Different formal
semantics are proposed to specify the behavior of timed Reo.
For example, in \cite{DBLP:conf/tase/Meng12}, a UTP-based
(\emph{Unifying Theories of Programming}) semantics is provided to verify connectors as
\emph{designs}. An operational semantics based on \emph{constraint automata} was raised by Baier et
al., where a variant of LTL was also proposed to describe the properties of timed connectors.

Formal verification techniques have also been proved applicable in timed connectors.
\cite{DBLP:conf/tase/LiCWS15} has shown us a comformance testing method on timed connectors.
Bounded model checking methods were also adapted in \cite{DBLP:journals/scp/Kemper12}, based on SAT
solvers. All these solutions seem practical and impressive, however, a common question is faced by
most of them, which is: \emph{How to obtain these formal models?}

@@

As a branch of \emph{machine learning} technique, active learning offers a way to obtain models from
low-level models. Works in \cite{DBLP:journals/mt/Daelemans10, DBLP:journals/iandc/Angluin87,
DBLP:conf/fase/RaffeltS06} shows impressive examples where active learning is used to extract
\emph{Mealy machines} or \emph{regular languages} without time domain.

@@

In this paper, we proposed a active learning framework (shown in Figure \ref{fig:howto}) to
automatically extract timed connectors from blackbox models. We present \emph{Parameterized Mealy
machine} as a parametierized semantics for timed Reo channels, which enables us to generate Mealy
machines with a given alphabet. Then the $L*$ algorithm is adapted and optimized to extract Mealy
machines with time action from the \emph{connectors in blackboxes}.

The rest of the paper is organized as follows. After this general introduction, in Section
\ref{sec:preliminaries} we briefly illustrate some basic concepts, including Reo the coordination
language, Mealy machine, and active automata learning. Section \ref{sec:semantics} defines the
Mealy-machine-based operational semantics of Reo, which is used in Section \ref{sec:activelearning}
to show how to extract Reo models from blackboxes by means of active learning. Finally, in Section
\ref{sec:experiment}, we discuss the optimization and implementation in \texttt{Golang}.

\section{Preliminaries} 
\label{sec:preliminaries}
\subsection{Reo Coordination Language} 
\label{sec:reo}
We provide here a brief overview of the main concepts of Reo, more details can be found in
\cite{DBLP:journals/mscs/Arbab04, DBLP:journals/scp/BaierSAR06}.

Reo is a channel based exogenous coordination language proposed by F. Arbab in
\cite{DBLP:journals/mscs/Arbab04}. 
A Reo model, also called \emph{connector}, provides the protocol
that formalizes the communication, synchronization and cooperation among the components which
communicate through the connector. Connectors can be defined with no knowledge of the components,
which makes Reo a powerful ``glue language'' in component-based
development\cite{DBLP:journals/sigsoft/Gill03}.

In Reo, complex connectors are made up of simpler ones, where the atomic connectors are called
\emph{channels}. Each channel has two \emph{channel ends}. There are two type of channel ends:
\emph{source} and \emph{sink}. Source channel ends accepts data into the channel, while sink
channels ends release them out of the channel.

@@

\begin{itemize}
  \item [-] A \emph{Sync} channel accepts a data item from its source end iff. the data item can be
    dispensed to its sink end simultaneously.
  \item [-] A \emph{LossySync} channel is always prepared to accept data items. These items will be
    send to its sink end simutaneously if possible, otherwise they will be dropped.
  \item [-] A \emph{SyncDrain} channel has two source ends and no sink end. It can accept a data
    item through one of its source end iff. a data item is also available for it to simultaneously
    accept through the other end. Then both two data items will be lost.
  \item [-] A \emph{FIFO1} channel is an asychronous channel with one buffer cell. It accepts a
    data item whenever the buffer is empty.  
\end{itemize}

@@

Components can be linked to source nodes or sink nodes. A component can write data items to its
corresponding source node only if the data item can be dispensed simultaneously to all source ends
on this node. Meanwhile, a component can read a data item only if there is at least one readable
sink end on its corresponding node. A mixed node non-determinstically selects and takes a data item
from one of its coincident sink channel ends and replicates it into all of its coincident source
channel ends.

@@

As shown in Figure \ref{fig:reoconnector}, we use a simple example to illustrate how complex
connectors are constructed and used in coordination. In this example, COM1,COM2 and
MONITOR are components. With an \emph{alternator} connector, MONITOR can receive data items from
COM1 and COM2 alternately. The \emph{alternator} example has been implemented in \texttt{Golang} as
shown in Section \ref{sec:reolib}.

@@

In this paper, we use \emph{Mealy machine} to model SUTs.
As an extension of \emph{finite state machine}, Mealy machine was first proposed by George. H. Mealy
in \cite{George1955A}. Compared with other variants, Mealy machines are designed to model
reactive systems, where outputs are determined not only by its current state but also the current
inputs. Besides, Mealy machines are supposed to be \emph{input enabled}, which means that all
possible inputs should be acceptable in all states. In other words, if an input is invalid for some
state, we need to manually use an additional state to describe such exceptions.

As far as we can see, various forms of Mealy machines are defined in different works, wherein some
are deterministic and some are not. Since active automata learning very much depends on the
system-under-learn to be deterministic, in this paper we formally define a deterministic version of
Mealy machine following \cite{DBLP:conf/sfm/SteffenHM11}.

@@

Active learning \cite{settles2010active} is a special case of semi-supervised machine learning where
a learning algorithm is able to interactively query the target systems to obtain the desired outputs
on certain inputs. With such flexibility, active learning makes it able to use targeted and efficient
queries to obtain more accurate models with smaller dataset.

@@

Figure \ref{fig:activelearning} shows the sketch of active automata learning, wherein:
\begin{itemize}
  \item[-] \emph{Teacher} and \emph{Student}. Active learning is an interactive process where
    students ask questions and teachers answer. Here learning algorithm plays the role of student.
  \item[-] \emph{Oracles} is an interface specifying which kind of questions can be answered by the
    teacher.
  \item[-] \emph{SUL} is an abbreviation of System Under Learn. The name comes from a well-known
    concept SUT (System Under Test) in software testing. In this case, we use blackbox models as our
    SULs.
  \item[-] \emph{CEGAR} indicates Counter-Example Guided Abstraction
    Refinement\cite{DBLP:conf/cav/ClarkeGJLV00}. In active learning, we need counter-examples to
    guide us on further quries and cover the undistinguished states.
\end{itemize}

When applying active automata learning on some model, firstly we assume that it should be equivalent
to some \emph{Mealy machine}. In other words, a deterministic model accepting a finite set of input
symbols and mapping them to a finite set of output symbols.

Such a model is encapsulated as a \emph{teacher} by the \emph{Oracle}, which handles
all communication with the model, both providing input and obtaining output. Also, it serves as a
so-called \emph{Minimal Adequate Teacher} interface, which is responsible for two types of
queries.

\begin{itemize}
  \item[-] \textbf{Membership Query} (hereinafter referred to as \emph{mq}) The name comes from
    some grammar-learning papers (e.g. \cite{DBLP:journals/iandc/Angluin87}), where \emph{mq} checks
    if a word is a member of certain language defined by the given grammar. When it comes to
    automata learning, \emph{mq} is supposed to provide \emph{simulation results} for given input
    sequences.
  \item[-] \textbf{Equivalence Query} (hereinaftr referred to as \emph{eq}) Given a hypothesis
    (usually constructed by the learning algorithm), \emph{eq} checks whether the hypothesis is
    equivalent to the system-under-learn and generates a counter-example if needed. Generally, 
    \emph{equivalence query} is irrealizable when SUT is a blackbox. So we tactically use membership
    queries to achieve the approximate results.
\end{itemize}

These queries are given by \emph{learning algorithms}, or so-called \emph{students} in Figure
\ref{fig:activelearning}. From the \emph{mq} results, a learning algorithm should construct a
\emph{hypothesis} and check it with \emph{eq}. If counter-examples are found, we turn back and
repeat the hypothesis construction until the equivalence query returns \emph{true}.

@@

\section{Timed Connectors as Mealy Machines}
\label{sec:semantics}
In this section, we show how a timed connector be transformed into a Mealy machine with time action.

Since time is not involed in original Mealy machine, we first discuss how to formalize the time
dimension in Mealy machine and timed Connectors. After that, we present a \emph{parameterized Mealy
machine} as a bridge between connectors and Mealy machines.

\subsection{Time Domain}
Time is involved in several extension version of Reo. For example, Timed
Reo\cite{DBLP:conf/sefm/ArbabBBR04}, Hybrid Reo\cite{DBLP:conf/icfem/ChenSS14}, etc.
Generally, these models are designed to handle real-time behavior where time is defined in
$\mathbb{R}$. Besides, we also found some works like \cite{DBLP:journals/fmsd/PrabhakarDM015} where
rational time indeed  makes things easier. In this paper, we choose the rational number field
$\mathbb{Q}$ as our time domain, which simplifies discretization of timed behaviors greatly.

@@

In real systems, the concept \emph{time precision} is widely used with the name ``clock-period''.
Most of the time, we know the clock-cycle of some hardware components, even without any idea of its
structure. With such precision $T$ given, it's reasonable to assume that all $t$-timers are actually
$nT$-timers. In following sections, we'll use $n$-timers instead.

Besides, we're going to add a ``T'' action in mealy machines. It indicates that a transition
will take a time unit to finish, and all outputs would come out after that.

\subsection{Parameterized Mealy Machine}
We present a model named \emph{parameterized Mealy machine} (hereinafter referred to as PMM)
to represent timed connectors. PMM is supposed to behave as a middle representation. Connectors are
defined as parameterized mealy machines, and composed via its production operator. Then original
mealy-machine model will be taken as semantics of PMM model.

Following the formal definition of Mealy machine in Section \ref{sec:preliminaries}, we present the
definition of \emph{parameterized Mealy machine} as follows.

\begin{definition}[Parameterized Mealy Machine]
  A \emph{Parameterized Mealy Machine} is defined as a 6-tuple $\mathcal{PM}=\langle
  S(\Sigma), s_0, I, O, \delta(\Sigma), \lambda(\Sigma)\rangle$ where 
  \begin{itemize}
    \item[-] $\Sigma$ is a \emph{finite} datum alphabet (hereinafter referred to as an alphabet)
    \item[-] $S$ is a function that maps an alphabet to a \emph{finite} set of
      states. We use $S(\Sigma)$ to denote the state set.
    \item[-] $I$ is a finite set of source-ends.
    \item[-] $O$ is a finite set of sink-ends.
    \item[-] $s_0$ is the initial state. It satisfies $\forall \Sigma,s_0\in S(\Sigma)$
    \item[-] $\delta$ maps a \emph{finite} datum alphabet to an \emph{output function}. We use
      $\delta(\Sigma):S(\Sigma)\times Input(\Sigma,I,O)\rightarrow Output(\Sigma,I,O)$ to denote the
      output function.
    \item[-] $\lambda$ maps an alphabet to a \emph{transition function}. We use
      $\lambda(\Sigma):S(\Sigma)\times Input(\Sigma,I,O)\rightarrow S(\Sigma)$ to denote the transiton
      function.
  \end{itemize}
\end{definition}

In the definition above, \emph{Input} and \emph{Output} are used to generate a set of input actions
and output actions from the corresponding alphabets and ends. The value of $Input(\Sigma,I,O)$ is
defined as a set of functions with on $I\cup O$ where $forall f\in Input(\Sigma,I,O)$, we have
\[
\forall i\in I, f(i)\in \Sigma\cup\{\bot\}\land \forall o\in
O,f(o)\in\{\rread, \rnoread\}
\]
where we use $\rread$ to indicate that a end-sink is ready for write, and $\rnoread$ otherwise. Note
that a $\bot$ means that there is \emph{no} data items on a source (or sink) end. In the same way,
$Output(\Sigma,I,O)$ is defined as a set of functions on domain $O$, with an additional symbol
$\rblock$ indicating an input failure. Similarly, $\forall f\in Output(\Sigma,I,O), f\neq \rblock$
implies
\[
\forall o\in O, f(o)\in \Sigma\cup\{\bot\}
\]

@@

If there no data item written to all sink ends, we use $\varnothing$ to denote the empty output. For
example, in the case given above, $\{B:\bot\}$ can be briefly rewritten as $\varnothing$.

@@

Now we can use Parameterized Mealy Machines to define a new semantics for timed Reo. Here we
take one asynchronous channel (FIFO1), one asynchronous channel (Sync) and one timed channel (Timer)
as examples.

@@

Besides, the concrete mealy machine (where $\Sigma=\{a\}$) is shown in Figure \ref{fig:pmmfifo},
where labels of edges are in form of $\frac{input}{output}$.
\begin{figure}[ht]
  \begin{center}
    \input{./images/pmmfifo.tex}
  \end{center}
  \caption{PMM-based Semantics of $\smap{FIFO1}_\Sigma$, where $\Sigma=\{a\}$}
  \label{fig:pmmfifo}
\end{figure}

@@

\begin{example}[PMM Semantics of n-Timer channel]
  Considering a n-Timer channel with a source-end A and a sink-end B, we define its PMM-based
  semantics as:
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_{i,d}|0\leq i\leq n, d\in \Sigma\}\cup\{q_0\}$
    \item[-] $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    \item[-] output function
      \begin{displaymath}
        \delta(\Sigma)(s,i)=\left\{
        \begin{array}[ht]{ll}
          (B:d) & s=q_{n,d}\land i=(A:d,B:\rread)\\
          \rblock & s=q_{n,d}\:\land \\
          & (i=T\lor i=(A:\_,B:\rnoread) \\
          \rblock & s=q_{j,d}\land 1 \leq j < n\land \\
          & (i=(A:\_,B:\rread)\lor \\
          & i=(A:d,B:\_))\\
          (B:\bot) & otherwise\\
        \end{array}
        \right.
      \end{displaymath}
    \item[-] transition function
      \begin{displaymath}
        \lambda(\Sigma)(s,i)=\left\{
        \begin{array}[ht]{ll}
          q_{j+1,d} & s=q_{j,d}\land i=T\land0 < j < n\\
          q_0 & s=q_{n,d}\land i=(A:\bot,B:\rread) \\
          q_{0,d} & s=q_{n,d'}\land i=(A:d,B:\rread) \\
          q_{0,d} & s=q_0 \land i=(A:d,B:\_) \\
          s & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{example}

@@

\todo{check the timed behavior}
\begin{definition}[Production Operator]
  Now we're going to define the production operator \emph{prod} of two parameterized Mealy machines as,
  \[
  prod(PM_1,PM_2)=PM_3
  \]
  as follows. Here we assume that $PM_2.O\cap PM_1.I=\varnothing$
  \begin{itemize}
  	\item[-] $\forall\Sigma, PM_3.S(\Sigma)=PM_1.S(\Sigma)\times PM_2.S(\Sigma)$
    \item[-] $PM_3.I=PM_1.I\cup PM_2.I-PM_1.O$
    \item[-] $PM_3.O=PM_1.O\cup PM_2.O-PM_2.I$
  \end{itemize}
  \emph{(Here we assume that sink ends of $PM_1$ can be connected to source ends $PM_2$, but not
  vise versa)}
  \begin{itemize}
    \item[-] $PM_3.s_0=(PM_1.s_0, PM_2.s_0)$
    \item[-] $\forall\Sigma, PM_3.\delta(\Sigma)((s_1,s_2), i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[ht]{ll}
          (Out_1 + Out_2)|_{PM_3.O} & Out_1\neq\rblock\land Out_2\neq\rblock \\
          \rblock & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
      where we have
      \begin{itemize}
        \item[*] $In_1 = i|_{PM_1.I}$
        \item[*] $Out_1 = PM_1.\delta(\Sigma)(s_1,In_1)$
        \item[*] $In_2 = (Out_1 + i)|_{PM_2.I}$
        \item[*] $Out_2 = PM_2.\delta(\Sigma)(s_2,In_2)$
      \end{itemize}
    \item[-] $\forall\Sigma, PM_3. \lambda(\Sigma)((s_1,s_2),i)=(s_1',s_2')$
      where we have
      \begin{itemize}
        \item[*] $s_1' = PM_1.\lambda(\Sigma)(s_1,In_1)$
        \item[*] $s_2' = PM_2.\lambda(\Sigma)(s_2,In_2)$
      \end{itemize}
  \end{itemize}
\end{definition}

The idea in \emph{production} is quite simple. The output of $PM_1$ would be provided as part of the
input of $PM_2$. Time actions T is only executed simutaneously betwenn $PM_1$ and $PM_2$.

As mentioned in the notation above, we cannot connect a sink end of $PM_2$ and a source end of
$PM_1$, which makes it difficult to construct connectors like alternator in Figure
\ref{fig:reoconnector}. Now we define a \emph{link} operation to connect sink ends and source ends
in the same connector.

\begin{definition}[Link Operator]
  The \emph{link} operation constructs connector by linking a sink end $OUT$ to a source end $IN$
  within the same connector.   
  \[
  PM' = link(PM, OUT, IN)
  \]
  Here we asuume that $IN\in PM.I$ and $OUT\in PM.O$.

>>>>

% finally no todos should exist in this draft
% \usepackage[textsize=tiny]{todonotes}

@@

Distributed real-time embedded system (DRES) is reforming our lives with the name \emph{IoT}, the
internet of things, wherein systems are usually composed of individual components and a
\emph{connector}. Such systems could be distributed logically or physically, which
makes coordination processes even more complicated. In this case, we need to specify these coordination
processes with so-called \emph{coordination languages}, so that formal techniques can be applied to
guarantee their reliability.

\emph{Timed Reo} is a real-time extension of the coordination language \emph{Reo}, by which
coordination process in DRES can be depicted clearly and intuitively. Different formal
semantics are proposed to specify the behavior of timed Reo.
For example, in \cite{DBLP:conf/tase/Meng12}, a UTP-based
(\emph{Unifying Theories of Programming}) semantics is provided to verify connectors as
\emph{designs}. An operational semantics based on \emph{constraint automata} was raised by Baier et
al.\cite{DBLP:conf/sefm/ArbabBBR04}, where a variant of LTL was also proposed to describe the properties of timed connectors.

Formal verification techniques have also been proved applicable in timed connectors, e.g.
\cite{DBLP:conf/tase/LiCWS15} has shown us a comformance testing method on timed connectors.
Bounded model checking methods based on SAT solvers were also adapted in
\cite{DBLP:journals/scp/Kemper12}.
All these solutions seem practical and impressive, however, a common question is faced by
most of them: \emph{how to obtain these formal models?}

@@

As a branch of \emph{machine learning} technique, active learning offers a way to obtain models from
low-level models. Works in \cite{DBLP:journals/mt/Daelemans10, DBLP:journals/iandc/Angluin87,
DBLP:conf/fase/RaffeltS06} shows interesting examples where active learning is used to extract
\emph{Mealy machines} or \emph{regular languages} without time domain.

@@

To address the problem, we proposed a active learning framework (shown in Figure \ref{fig:howto}) to
automatically extract timed connectors from blackbox models. To achieve this goal, firstly we
present \emph{parameterized Mealy machine} as a parametierized semantics for timed Reo channels,
which is able to generate concrete Mealy machines with a given alphabet. Then the $L*$ algorithm
\cite{DBLP:journals/iandc/Angluin87}
is adapted and optimized to extract Mealy machines with time action from the \emph{connectors in
blackboxes}.

The rest of the paper is organized as follows. After this general introduction, in Section
\ref{sec:preliminaries} we briefly illustrate some basic concepts, including Reo the coordination
language, Mealy machine, and active automata learning. Section \ref{sec:semantics} defines an
operational semantics of Reo, which is used in Section \ref{sec:activelearning} to show how to
extract Reo models from blackboxes by means of active learning. Finally, in Section
\ref{sec:experiment}, we discuss the implementation and optimization in \texttt{Golang}.

\section{Preliminaries} 
\label{sec:preliminaries}
\subsection{Reo Coordination Language} 
\label{sec:reo}
We provide here a brief overview of the main concepts in Reo. Further details can be found in
\cite{DBLP:journals/mscs/Arbab04, DBLP:journals/scp/BaierSAR06}.

Reo is a channel based exogenous coordination language proposed by F. Arbab in
\cite{DBLP:journals/mscs/Arbab04}. 
A Reo model, also called \emph{connector}, provides the protocol
that formalizes the communication, synchronization and cooperation among the components which
communicate through the connector. Connectors can be defined without dependence on components,
which makes Reo a powerful ``glue language'' in component-based
development\cite{DBLP:journals/sigsoft/Gill03}.

In Reo, complex connectors are made up of simpler ones. The atomic connectors are called
\emph{channels}, and each of them has two \emph{channel ends}. There are two type of channel ends:
\emph{source} and \emph{sink}. Source channel ends accepts data into the channel, while sink
channels ends dispense them out of the channel.

@@

\begin{itemize}
  \item [-] A \emph{Sync} channel accepts a data item from its source end iff the data item can be
    dispensed to its sink end simultaneously.
  \item [-] A \emph{LossySync} channel is always prepared to accept data items. These items will be
    send to its sink end simutaneously if possible, otherwise they will be dropped.
  \item [-] A \emph{SyncDrain} channel has two source ends and no sink end. It can accept a data
    item through one of its source end iff a data item is also available to accept simultaneously
    through the other end. Then both items will be dropped.
  \item [-] A \emph{FIFO1} channel is an asychronous channel with a buffer cell. It accepts a
    data item whenever the buffer is empty. 
\end{itemize}

@@

Components can be linked to source nodes or sink nodes. A component can write data items to its
corresponding source node only if the data item can be dispensed simultaneously to all source ends
on this node. Similarly, a component can read a data item only if there is at least one readable
sink end on its corresponding node. A mixed node non-determinstically selects and takes a data item
from one of its coincident sink ends and replicates it into all of its coincident source ends.

@@

As shown in Figure \ref{fig:reoconnector}, we use a simple example to illustrate how complex
connectors are constructed and used in coordination. In this example, COM1,COM2 and
MONITOR are components. With an \emph{alternator} connector, MONITOR can obtain data items from
COM1 and COM2 alternately. The \emph{alternator} example has been implemented in \texttt{Golang} as
shown in Section \ref{sec:reolib}.
Besides, all the connectors are supposed to be \emph{deterministic} hereinafter to meet the
requirement of active learning algorithm L*.

@@

In this work, we use \emph{Mealy machines} to model Reo connectors.
As an extension of \emph{finite state machine}, Mealy machine was first proposed by George. H. Mealy
in \cite{George1955A}. Compared with other variants, Mealy machines are designed to model
reactive systems, where outputs are determined not only by its current state but also the current
inputs. Besides, Mealy machines are supposed to be \emph{input enabled}, which means that all
possible inputs should be acceptable in all states. In other words, if an input is invalid for some
state, we need to manually define an additional state to describe such exceptions.

Various forms of Mealy machines are defined in different works. 
Since active automata learning very much depends on the system-under-learn to be deterministic, in
this paper we formally define a deterministic Mealy machine following
\cite{DBLP:conf/sfm/SteffenHM11}.

@@

Active learning \cite{settles2010active} is a special case of semi-supervised machine learning where
a learning algorithm is able to interactively query the target systems to obtain the desired outputs
under certain inputs. By well-designed query strategy, active learning is able to obtain more
accurate models with smaller dataset.

@@

Figure \ref{fig:activelearning} shows the sketch of active learning, wherein:
\begin{itemize}
  \item[-] \emph{Teacher} and \emph{Student}: Active learning is an interactive process where
    students ask questions and teachers answer. Here learning algorithm plays the role of student.
  \item[-] \emph{Oracles} is an interface specifying which kind of questions can be answered by the
    teacher.
  \item[-] \emph{SUL} is an abbreviation of System Under Learn. The name comes from a well-known
    concept SUT (System Under Test) in software testing. In this case, we take blackbox models as our
    SULs.
  \item[-] \emph{CEGAR} indicates Counter-Example Guided Abstraction
    Refinement\cite{DBLP:conf/cav/ClarkeGJLV00}. In active learning, we need counter-examples to
    guide us on further quries and cover the undistinguished states.
\end{itemize}

When applying active automata learning on some model, firstly we assume that it should be equivalent
to some \emph{Mealy machine}, in other words, a deterministic model accepting a finite set of input
symbols and mapping them to a finite set of output symbols.

Such a model is encapsulated as a \emph{teacher} by the \emph{Oracle}, which handles
all communication with the model. The oracle also serves as a so-called \emph{Minimal Adequate Teacher}
interface, which is responsible for two types of queries.

\begin{itemize}
  \item[-] \textbf{Membership Query} (hereinafter referred to as \emph{mq}) The name comes from
    some grammar-learning papers (e.g. \cite{DBLP:journals/iandc/Angluin87}), where \emph{mq} checks
    if a word is a member of certain language defined by the given grammar. When it comes to
    automata learning, \emph{mq} is supposed to provide \emph{simulation results} for given input
    sequences.
  \item[-] \textbf{Equivalence Query} (hereinaftr referred to as \emph{eq}) Given a hypothesis
    (usually constructed by the learning algorithm), \emph{eq} checks whether the hypothesis is
    equivalent to the system-under-learn and generates a counter-example if needed. Generally, 
    \emph{equivalence query} is irrealizable when SUL is a blackbox. So we tactically use membership
    queries to achieve the approximate results.
\end{itemize}

These queries are given by \emph{learning algorithms}, or so-called \emph{students} in Figure
\ref{fig:activelearning}. From the \emph{mq} results, a learning algorithm constructs a
\emph{hypothesis} and then check it with \emph{eq}. If counter-examples are found, we turn back and
repeat the hypothesis construction until the equivalence query returns \emph{true}.

@@

\section{Timed Connectors as Mealy Machines}
\label{sec:semantics}
In this section, we illustrate how to transform a timed connector into a Mealy machine with time
action.
Since time is not involed in original Mealy machine, we first discuss how to formalize the time
dimension in Mealy machine. After that, we present a \emph{parameterized Mealy
machine} as a bridge between connectors and Mealy machines.

\subsection{Time Domain}
Time is involved in several extensions of Reo. For example, timed
Reo\cite{DBLP:conf/sefm/ArbabBBR04}, hybrid Reo\cite{DBLP:conf/icfem/ChenSS14}, etc.
Generally, these models are designed to handle real-time behavior where time is defined in
$\mathbb{R}$. However, we also found some works like \cite{DBLP:journals/fmsd/PrabhakarDM015} where
rational time indeed  makes things easier. In this paper, we choose the rational number field
$\mathbb{Q}$ as our time domain, which simplifies discretization of timed behaviors greatly.

@@

In real systems, the concept \emph{time precision} is well known with other names like ``clock-period''.
Most of the time, we know the clock-period of some hardware components, even without any idea of its
structure. With such precision $dt$ given, it's reasonable to assume that all $t$-timers are actually
$ndt$-timers. In following sections, we'll use $n$-Timers instead.

Besides, we're going to add a ``T'' action in Mealy machines. It indicates that a transition
will take a time unit to finish.

\subsection{Parameterized Mealy Machine}
We present a model named \emph{parameterized Mealy machine} (hereinafter referred to as PMM)
to model timed connectors. PMM is supposed to behave as a middle representation between Mealy
machines and timed connectors. Connectors are firstly defined as parameterized Mealy machines, and
composed by production operator and link operator. Then, concrete Mealy-machine model can be
generated from the PMM model.
Following the formal definition of Mealy machine in Section \ref{sec:preliminaries}, we define the
\emph{parameterized Mealy machine} as follows.

\begin{definition}[Parameterized Mealy Machine]
  A \emph{Parameterized Mealy Machine} is defined as a 6-tuple $\mathcal{PM}=\langle
  S, s_0, I, O, \delta, \lambda\rangle$ with a parameter named $\Sigma$ where 
  \begin{itemize}
    \item[-] $\Sigma$, the parameter, is a \emph{finite} datum alphabet
    \item[-] $S$ is a function that maps an alphabet to a \emph{finite} set of
      states. We use $S(\Sigma)$ to denote the state set.
    \item[-] $I$ is a finite set of source-ends.
    \item[-] $O$ is a finite set of sink-ends.
    \item[-] $s_0$ is the initial state. It satisfies $\forall \Sigma,s_0\in S(\Sigma)$
    \item[-] $\delta$ maps a \emph{finite} datum alphabet to an \emph{output function}. We use
      $\delta(\Sigma):S(\Sigma)\times Input(\Sigma,I,O)\rightarrow Output(\Sigma,I,O)$ to denote the
      output function.
    \item[-] $\lambda$ maps an alphabet to a \emph{transition function}. We use
      $\lambda(\Sigma):S(\Sigma)\times Input(\Sigma,I,O)\rightarrow S(\Sigma)$ to denote the transiton
      function.
  \end{itemize}
\end{definition}

In the definition above, \emph{Input} and \emph{Output} are used to generate input actions
and output actions from the corresponding alphabets and source/sink ends. The value of $Input(\Sigma,I,O)$ is
defined as a set of functions on $I\cup O$ where $\forall f\in Input(\Sigma,I,O)$ and an additional
\emph{time action} T, we have
\[
\forall i\in I\backslash\{T\}, f(i)\in \Sigma\cup\{\bot\}\land \forall o\in
O,f(o)\in\{\rread, \rnoread\}
\]
where we use $\rread$ to indicate that an end-sink is ready for write, and $\rnoread$ otherwise. Note
that a $\bot$ means that there is \emph{no} data items on a source (or sink) end. In the same way,
$Output(\Sigma,I,O)$ is defined as a set of functions on domain $O$, with an additional symbol
$\rblock$ indicating an input failure. Similarly, $\forall f\in Output(\Sigma,I,O), f\neq \rblock$
implies
\[
\forall o\in O, f(o)\in \Sigma\cup\{\bot\}
\]

@@

If there is no data item written to any sink ends, we use $\varnothing$ to denote such an empty
output. For example, in the case given above, $\{B:\bot\}$ can be briefly rewritten as $\varnothing$.

@@

Now we can use parameterized Mealy Machines to define a new semantics for timed Reo. Here we
take one asynchronous channel (FIFO1), one asynchronous channel (Sync) and one timed channel (Timer)
as examples.

@@

Besides, the concrete Mealy machine (where $\Sigma=\{a\}$) is shown in Figure \ref{fig:pmmfifo},
where labels of edges are in form of $\frac{input}{output}$.
\begin{figure}[ht]
  \begin{center}
    \input{./images/pmmfifo.tex}
  \end{center}
  \caption{PMM-based Semantics of $\smap{FIFO1}_\Sigma$, where $\Sigma=\{a\}$}
  \label{fig:pmmfifo}
\end{figure}

@@

\begin{example}[PMM Semantics of n-Timer channel]
  Considering an n-Timer channel with a source-end A and a sink-end B, we define its PMM-based
  semantics as:
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_{i,d}|0\leq i\leq n, d\in \Sigma\}\cup\{q_0\}$
    \item[-] $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    \item[-] output function
      \begin{displaymath}
        \delta(\Sigma)(s,i)=\left\{
        \begin{array}[ht]{ll}
          (B:d) & s=q_{n,d}\land i=(A:d,B:\rread)\\
          \rblock & s=q_{n,d}\:\land \\
          & (i=T\lor i=(A:\_,B:\rnoread) \\
          \rblock & s=q_{j,d}\land 1 \leq j < n\land \\
          & (i=(A:\_,B:\rread)\lor \\
          & i=(A:d,B:\_))\\
          (B:\bot) & otherwise\\
        \end{array}
        \right.
      \end{displaymath}
    \item[-] transition function
      \begin{displaymath}
        \lambda(\Sigma)(s,i)=\left\{
        \begin{array}[ht]{ll}
          q_{j+1,d} & s=q_{j,d}\land i=T\land0 < j < n\\
          q_0 & s=q_{n,d}\land i=(A:\bot,B:\rread) \\
          q_{0,d} & s=q_{n,d'}\land i=(A:d,B:\rread) \\
          q_{0,d} & s=q_0 \land i=(A:d,B:\_) \\
          s & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{example}

@@

\begin{definition}[Production Operator]
  Now we're going to define the production operator \emph{prod} of two parameterized Mealy machines as,
  \[
  prod(PM_1,PM_2)=PM_3
  \]
  as follows. Here we assume that $PM_2.O\cap PM_1.I=\varnothing$
  \begin{itemize}
  	\item[-] $\forall\Sigma, PM_3.S(\Sigma)=PM_1.S(\Sigma)\times PM_2.S(\Sigma)$
    \item[-] $PM_3.I=PM_1.I\cup PM_2.I-PM_1.O$
    \item[-] $PM_3.O=PM_1.O\cup PM_2.O-PM_2.I$
  \end{itemize}
  \emph{(Here we assume that sink ends of $PM_1$ can be connected to source ends $PM_2$, but not
  vise versa)}
  \begin{itemize}
    \item[-] $PM_3.s_0=(PM_1.s_0, PM_2.s_0)$
    \item[-] $\forall\Sigma, PM_3.\delta(\Sigma)((s_1,s_2), i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[ht]{ll}
          (Out_1 + Out_2)\downarrow_{PM_3.O} & Out_1\neq\rblock\land Out_2\neq\rblock \\
          \rempty & i = T \\
          \rblock & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
      where we have
      \begin{itemize}
        \item[*] $In_1 = i\downarrow_{PM_1.I}$
        \item[*] $Out_1 = PM_1.\delta(\Sigma)(s_1,In_1)$
        \item[*] $In_2 = (Out_1 + i)\downarrow_{PM_2.I}$
        \item[*] $Out_2 = PM_2.\delta(\Sigma)(s_2,In_2)$
      \end{itemize}
    \item[-] $\forall\Sigma, PM_3. \lambda(\Sigma)((s_1,s_2),i)=(s_1',s_2')$
      where we have
      \begin{itemize}
        \item[*] $s_1' = PM_1.\lambda(\Sigma)(s_1,In_1)$
        \item[*] $s_2' = PM_2.\lambda(\Sigma)(s_2,In_2)$
      \end{itemize}
  \end{itemize}
\end{definition}

The idea in \emph{production} is quite simple. The output of $PM_1$ would be provided as part of the
input of $PM_2$. Time actions T is only executed simutaneously between $PM_1$ and $PM_2$.

As mentioned in the notation above, we cannot connect a sink end of $PM_2$ to a source end of
$PM_1$, which makes it difficult to construct connectors like alternator in Figure
\ref{fig:reoconnector}. To address the problem, we define a \emph{link} operation to connect sink
ends and source ends in the same connector.

\begin{definition}[Link Operator]
  The \emph{link} operation constructs connector by connecting a sink end $OUT$ to a source end $IN$
  within the same connector.   
  \[
  PM' = link(PM, OUT, IN)
  \]
  Here we asuume that $IN\in PM.I$ and $OUT\in PM.O$.
